---
title: "Penguin Flipper Lengths ANOVA"
author: "Alex Ivina"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###1. Load packages and read in the penguins.csv data
```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Our question: Is there a significant difference in mean flipper lenghths for female Chinstrap, Gentoo, and Adeline Penguins near Palmer Archipelago?

Maybe I can use a one-way ANOVA....
- Independent observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

```{r}
counts <- with(penguins, table(Species, Sex))

penguins2 <- penguins %>% 
  mutate(
    common_name = 
      case_when(
        Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", 
        Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adeline"
      )
  ) %>% 
  select(-Species) %>% 
  filter(Sex == "FEMALE")

```

 Exploratory graphs and Levene's test for equal variances
 
```{r}
hists <- ggplot(penguins2, aes(x = `Flipper Length (mm)`)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale ="free")

hists

qqs <- ggplot(penguins2, aes(sample = `Flipper Length (mm)`)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~ common_name, scale = "free")

qqs

# Based on qq-plots and histogram, observations for each group appear normally distrubuted. Also, n > 30 for all groups, so means will be normally distrubited by CLT.

```
 
 Evaluate variances and perform a Levene's Test for equal variances
 
```{r}
# If the largest sample variances is < 4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variances = var(`Flipper Length (mm)`))
    
variances # The largest variances (33.1) is definitely not more than 4x greater than the smallest variances (15.2)

# Levene's Test
# leveneTest function in the 'car' package

levene_penguin <-leveneTest(`Flipper Length (mm)` ~ common_name, data = penguins2)
levene_penguin

# H0: No differences in variances (variances are equal)
# HA: Variances are NOT equal

# There is no significant differences in variances across species.


```
 
 One-way ANOVA for penguin flipper lengths by Species (common_name)
 
```{r}

# No changes these flipper lengths can be taken from the population of the same fipper lengths

# H0: Means across all groups are equal
# HA: At least two group means differe significantly

penguin_aov <- aov(`Flipper Length (mm)` ~ common_name, data = penguins2)
summary(penguin_aov)

# Post-hoc testing by Turkey's HSD
# H0: No significant difference in pairwise means
# HA: There is a significant difference in pairwise means

penguins_ph <- TukeyHSD(penguin_aov)
penguins_ph

# aov.test function in the 'onewaytests' package DOES allow  for inline references 


```
 
 Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).
 